Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2017-10-26 LaTeX/Turnover.tex

% from Bai et al. (2016) ``[Segregation distortion] phenomenon is commonly detected via genetic mapping and has been documented in various species, including mouse (Eversley et al., 2010; Casellas et al., 2012), Drosophila (Phadnis and Orr, 2009; Larracuente and Presgraves, 2012; McDermott and Noor, 2012), Tigriopus (Pritchard et al., 2011), rice (Koide et al., 2012; Reflinur et al., 2014; Xu et al., 2014), maize (Tang et al., 2013), and cotton (Yu et al., 2011; Hulse-Kemp et al., 2015).'' Also more in that paper that might be useful. 
%Drosophila http://www.biorxiv.org/content/early/2017/02/06/106336
\textcolor{red}{Genetic mapping experiments, which are typically designed to minimize selection in diploids, have revealed segregation distortion in various species, including mice, Drosophila, Rice, Maize, Wheat, Barley, Cotton... 
In some of these cases, biased segregation has been attributed to meiotic drive and/or gametic selection \textcolor{red}{(Leppala et al. 2013, Didion et al. 2015, 2016 Xu et al 2013 (rice), Fishman...)}. } 
Meiotic drive generally occurs either during the production of male or female gametes only \citep{Ubeda:2005gw,Lindholm:2016cw}.
Gametic competition is also typically sex specific, occurring primarily among male gametes, because there are typically many more pollen/sperm than required for fertilization.
Gametic competition may be particularly common in plants, in which 60-70\% of all genes are expressed in the male gametophyte and these genes exhibit stronger signatures of selection than random genes \citep{Borg:2009jpa,Arunkumar:2013iq,Gossmann:2014dua}.
In addition, artificial selection pressures applied to male gametophytes are known to cause a response to selection \citep[e.g.,][]{Hormaza:1996cv,Ravikumar:2003uo,Hedhly:2004iv,Clarke:2004ir}. 
A smaller proportion of genes are thought to be expressed and selected during competition in animal sperm, although precise estimates are uncertain \citep{Zheng:2001fi,JOSEPH:2004haa,Vibranovski:2010et}. 
Recent studies have demonstrated that sperm competition in animals can alter haploid allele frequencies and increase offspring fitness \citep{Immler:2014im,Alavioon2017}.

There are various ways in which a period of haploid selection could influence transitions between sex-determination systems. 
If we assume that haploid selection at any particular locus predominantly occurs in one sex (e.g., meiotic drive during spermatogenesis), then such loci experience a form of sex-specific selection. 
In this respect, we might expect that haploid selection would affect transitions between sex-determination systems in a similar manner to sex-specific diploid selection \citep[as explored by][]{vanDoorn:2007eu,vanDoorn:2010hu}. 
That is, new masculinizing mutations (neo-Y chromosomes) could be favoured via associations with alleles that are beneficial in the male haploid stage. 
On the other hand, sex ratios can also become biased by linkage between the sex-determining region and a locus that harbours genetic variation in haploid fitness. 
For example, there are several known cases of sex-ratio bias caused by sex-linked meiotic drive alleles \citep[][Chapter 3]{Burt:2006} or selection among X- and Y-bearing pollen \citep{Lloyd:1974tz,Conn:1981uw,Stehlik:2005ul,Stehlik:2006to,Field:2012fd,Field:2013cc}. 
It is not immediately clear how the spread of new sex-determination systems would be influenced by the combination of sex-ratio biases and associations between haploid selected loci and sex-determining regions.

@@

We find that sex-ratio biases caused by haploid selection can exert Fisherian sex-ratio selection upon novel sex-determiners but that their spread is also determined by selection on genetically-associated alleles. 
%Indeed, it is only when haploid-selected loci are tightly linked to the ancestral sex-determining region (and so sex-ratio biases are initially large) that we see an asymmetry between selection for XY to ZW transitions and ZW to XY transitions (e.g., because haploid selection in males only causes biased zygotic sex ratios in an ancestrally XY system).
Consequently, Fisherian sex ratio selection does not dominate and it is possible for selection on linked alleles to drive turnover between sex-determining systems despite causing increases in sex-ratio bias. 
In addition to considering haploid selection, another novel development in our model is that we consider loci that are in very tight linkage with the ancestral sex-determining region.
We show that transitions between male and female heterogamety can then evolve despite the fact that the neo-sex-determining locus is less closely linked to a locus under selection and therefore disrupts favourable ancestral associations between sex and the alleles selected in that sex.
%Such transitions are not favoured in models lacking tight linkage and/or haploid selection. 
%Secondly, we allow sex-specific haploid selection to occur on a locus in tight or loose linkage with the ancestral sex-determining region. 
%In addition, when linkage between the ancestral sex-determining locus and a locus under haploid selection is tight, sex ratio biases can be strong such that there is an asymmetry between invasion of ancestrally XY and ancestrally ZW systems because, e.g., haploid selection in males only causes biased zygotic sex ratios in an ancestrally XY system.

@@

%\textcolor{blue}{
Neo-W alleles, on the other hand, can invade an ancestral XY system under some conditions (the full invasion conditions are given in the appendix; equations \ref{Binvasion} and \ref{Ainvasion}). 
That is, selection on loci within the non-recombining region of the SDR can favour the invasion of a less closely linked neo-W (Figure \ref{fig:SexAntagTighter}). 
%Although haploid selection can favour neo-W alleles because the ancestral sex ratio becomes male biased, this is not the only circumstance in which less tightly linked neo-W alleles invade. 
%\textcolor{red}{Sally: unpack, give one condition and relate to figures}
%For example, we note that it is possible for both neo-W haplotypes to spread ($\lambda_{mA}>1$ and $\lambda_{ma}>1$), in which case neo-W invasion can occur regardless of linkage to the selected locus. 
%For example, unlinked neo-W alleles can invade in the absence of any haploid selection. 
%This result is unexpected given the results of \cite{vanDoorn:2010hu}, who did not explicitly calculate equilibrium allele frequencies under tight linkage and generally concluded that heterogametic transitions can only occur when neo-sex-determining alleles are in tighter linkage with loci under sex-specific diploid selection. 
In fact, with tight linkage between the ancestral SDR and the selected locus, haploid selection and/or overdominance can favour completely unlinked neo-W alleles ($R=1/2$), allowing autosomes to become new sex chromosomes.
To develop an intuition for how less closely linked neo-W alleles invade, we first focus on cases where there is no haploid selection and then discuss the additional effect of haploid selection (more details in the appendix). 
%}
%Briefly, a neo-W can spread when the females it produces have higher fitness than the females produced under the ancestral SDR, which all have genotype XX. X chromosomes are sometimes found in males and therefore allele frequencies on the X can be suboptimal for females. In addition, the neo-W produces females from matings with Y-bearing male gametes. The allele associated with the Y can also be beneficial in females. 
%the neo-W invades under the ancestral XY sex-determining system, the X chromosome may not be able to as specialised on a female-beneficial allele

@@

%Haploid selection impacts the spread of neo-W haplotypes through its direct selective effect on females and their gametes, its indirect effect on the marginal fitness of alleles in females resulting from haploid selection in male gametes, and its effect on the sex ratio.
%These impacts expand the scenarios under which neo-W haplotypes can spread.
%For instance, overdominance is no longer required for neo-W-$a$ haplotypes to spread when there is haploid selection (e.g., when $a$ is female-beneficial, $A$ is male-beneficial, and meiotic drive in males favours $a$, Figure \ref{fig:regionMaleDrive}B).
\textcolor{blue}{haploid selection?}

@@

Significantly, equation \eqref{eq:lambda_ESD} is the same whether ESD is invading an ancestrally XY or ZW system (because $\lambda_{Y',XY} = \lambda_{W',ZW}$ and $\lambda_{W',XY} = \lambda_{Y',ZW}$).
Thus, because the sex ratio is only biased by male haploid selection when the ancestral sex-determination system is XY, Fisherian sex-ratio selection alone does not explain the invasion of ESD under weak selection. 
Specifically, with male haploid selection, the neo-ESD is equally likely to invade when it equalizes the zygotic sex ratio (through $\lambda_{W',XY}$) and when it doesn't (through $\lambda_{Y',ZW}$). 
%This is again because haploid selection can facilitate the turnover of sex chromosomes either through sex ratio biases or through associations between the new sex determining region with selected alleles. 
%With weak selection, these two forces are nearly equal in strength. 
In addition, we note that ESD may not invade, even if the sex ratio is initially biased (e.g., with drive in males only, $r<1/2$, $h^\female=h^\male$, and $s^\female s^\male<0$, such that $\lambda_{W',XY}<1$, see Table \ref{tab:specialcases}).

@@

\textcolor{red}{check last condition and the stability condition below are correct}
\textcolor{blue}{The last condition looks good to me, although in your Turnover-norec-MFS.nb you look at YA fixed, so you have to flip everything (so I made Turnover-norec-MFS-MMO.nb to do this).
The one issue I can find here is that you can also prevent $\lambda>1$ when the slope and intercept of the quadratic at $\lambda=1$ are negative (you only looked at both being positive). In this case we need $4w_{A}^{\female}w_{AA}^{\female}<(1-\alpha_\Delta^{\female})w_{a}^{\female}w_{Aa}^{\female}$, which also prevents $\psi>0$.
I've added this in.
It could also be the case that the slope and intercept are the same sign but the roots are imaginary - but this is never the case here.
Stability condition below looks good to me (from matt version of turnoverSOM-MIKE.nb).}

@@

Still considering $w_{Aa}^\female>w_{AA}^\female$, the neo-W can also spread alongside the $a$ allele ($\lambda_{ma}>1$) if $w_{aa}^\female$ is large enough such that $(w_{Aa}^\female+w_{aa}^\female)/2>w_{AA}^\female$.
%there is sufficiently strong underdominance in females ($w_{aa}^\female>w_{Aa}^\female$) \textcolor{blue}{[this is describing directional selection, not overdominance - check the conditions for the case we want to talk about]}, such that $(w_{Aa}^\female+w_{aa}^\female)/2>w_{AA}^\female$.  
This can occur with overdominance or directional selection for $a$ in females (Figure \ref{fig:regionplots}B,C).
In this case, $a$ is not favoured enough in females near the equilibrium (comparing $Aa$ to $AA$ genotypes) to prevent $A$ from stably fixing on the X, and yet the neo-W can spread with $a$ because it produces female $aa$ individuals by capturing Y-$a$ haplotypes.
\textcolor{blue}{mike, could you check this paragraph? i fixed an old problem, and want to make sure i haven't introduced any new errors}

@@

\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=1.5\linewidth]{PositionPlot_SexAntagTighter_MaleDrive_Mike}}
\caption{
When there is sexually-antagonistic selection and haploid selection, a neo-W may invade for any $R$.
\textcolor{red}{
Check that we mention Sally's result that invasion cannot occur with sexually-antagonistic selection and $R=1/2$ (currently only in legend for figure \ref{fig:SexAntagTighter}). 
%Add asterisk to Figure \ref{fig:regionMaleDrive}A to show parameters used in this plot. 
}
\textcolor{blue}{This is now discussed in a round-about way at the beginning of the tight linkage invasion section in the main text. }
Parameters given by asterisk in Figure \ref{fig:regionMaleDrive}A: $w_{AA}^\female=1.05$, $w_{Aa}^\Hermaphrodite=1$, $w_{aa}^\female=0.85$, $w_{AA}^\male=0.85$, $w_{aa}^\male=1.05$, $\alpha_{\Delta}^\male=-0.08$.
}
\label{fig:SexAntagTighterMaleDrive}
\end{figure}

\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=\linewidth]{Region_Plot_combined_MaleDrive_Mike}}
\caption{
ABC, $\alpha_{\Delta}^\male=-0.08$
DEF, $\alpha_{\Delta}^\male=0.08$. 
%\textcolor{red}{Panel F mislabelled, should have $\lambda_{Wa}>1$ \& $\lambda_{WA}<1$ as the upper label that has the line.}
}
\label{fig:regionMaleDrive}
\end{figure}

\newpage
\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=\linewidth]{Region_Plot_combined_MaleGS_Mike}}
\caption{
ABC, $w_{a}^\male=1.16$, $w_{A}^\male=1$.
DEF, $w_{a}^\male=1$, $w_{A}^\male=1.16$. 
\textcolor{red}{I thought that re-running this with $w_{a}^\male=1.16$, $w_{A}^\male=0.84$ and $w_{a}^\male=0.84$, $w_{A}^\male=1.16$ will mean that it matches exactly with the meiotic drive example. Checking for panel A, the result was qualitatively similar (e.g., region where both $\lambda$'s are greater than one in panel A, as you might expect from equation \eqref{Binvasion}, however the region where a polymorphism is stable is also altered so they don't exactly match. }
\textcolor{blue}{trouble is selection only in heterozygotes with drive?}
}
\label{fig:regionMaleGS}
\end{figure}

\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=\linewidth]{Region_Plot_combined_FemaleDrive_Mike}}
\caption{
ABC, $\alpha_{\Delta}^\female=-0.16$
DEF, $\alpha_{\Delta}^\female=0.16$. $\lambda$s are increased for the haplotype that is favoured by female haploid selection, the stability conditions are also affected. 
}
\label{fig:regionFemaleDrive}
\end{figure}

\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=\linewidth]{Region_Plot_combined_FemaleGS_Mike}}
\caption{
ABC, $w_{a}^\female=1.16$, $w_{A}^\female=1$.
DEF, $w_{a}^\female=1$, $w_{A}^\female=1.16$. 
%\textcolor{red}{Panel F mislabelled, should have $\lambda_{Wa}>1$ \& $\lambda_{WA}<1$ as the upper label that has the line.}
}
\label{fig:regionFemaleGS}
\end{figure}

@@

\begin{landscape}
\begin{figure}[!h]
\centering
\centerline{
\includegraphics[width=\linewidth]{Freq_plot_combined_PloidAntag_Matt}
}
\caption{
Dynamics of all sex-determining alleles in each sex (males dashed).
Red is neo-W, Blue is Y, Black is $A$.
Panel A has the same parameters as the green curve in Figure \ref{fig:SexRatioBad}B (ploidy-antagonism with male drive).
Panel B has the same parameters as the green curve in Figure \ref{fig:temporalOverdominance}C (overdominance in both sexes).
Panel C has the same parameters as the green curve in Figure \ref{fig:regionMaleDrive}C (sexual-antagonism with male drive).
Panel D has the same parameters as the red curve in Figure \ref{fig:regionPloidAntag}F, except $r=0$ (ploidy-antagonism with pollen competition).
Panel A shows complete sex-determination turnover (XY $\rightarrow$ ZW) whereas panels B-D show the evolution of polymorphic sex-determining systems (X and Y still segregating and the neo-W has a frequency in females that is less than 1/2).
All begin from equilibrium that would have $a$ fixed on the Y if $r=0$.
All maintain a polymorphism at $\mathbf{A}$ locus.
}
\label{fig:freqAll}
\end{figure}
\end{landscape}

\begin{figure}[!h]
\centering
\centerline{
\includegraphics[width=\linewidth]{Sallys_R_plot.pdf}
}
\caption{
\textcolor{blue}{is this the one?}
}
\label{fig:lambda_R}
\end{figure}

\newpage
\textcolor{red}{
Add Sally's figure showing lambda for small r near equil A versus near equil B. Add references to this figure to appendix where we discuss whether lambdas can be greater than 1 with sexually antagonistic selection.}
\textcolor{blue}{not sure which one you are talking about, but see Figure \ref{fig:lambda_R} }

\textcolor{red}{
Perhaps it would also be useful to add an 8 panel figure that features ploidally antagonistic selection. For each type of haploid selection (gametic competition/ meiotic drive in males/females), give a regionplot where $h^\male=h^\female$, e.g., $h^\male=h^\female=0.75$ (or perhaps the value of $h$ we use in the regionplots we have, in which $w_{aa}=0.85$, $w_{Aa}=1$, $w_{AA}=1.05$). 
Matt made a figure like this before but both Y$a$ and Y$A$ equilibria were plotted and there was no outline showing where the Y$a$ equilibrium is stable (as in Figure \ref{fig:regionplots}). In Matts plot the axes were $s^\Hermaphrodite$ and $\alpha_{\Delta}^\Hermaphrodite$. 
Add an asterisk to each region plot and show invasion in another panel, using those parameters and various $R$ (e.g., in the stye of \ref{fig:temporalOverdominance}). 
In an email, Sally has an example of ploidally-antagonistic selection where the neo-W fixes and $R=1/2$. This would cover that case and more. }
\textcolor{blue}{
made an attempt (Figure \ref{fig:regionPloidAntag})
}

>>>>

% from Bai et al. (2016) ``[Segregation distortion] phenomenon is commonly detected via genetic mapping and has been documented in various species, including mouse (Eversley et al., 2010; Casellas et al., 2012), Drosophila (Phadnis and Orr, 2009; Larracuente and Presgraves, 2012; McDermott and Noor, 2012), Tigriopus (Pritchard et al., 2011), rice (Koide et al., 2012; Reflinur et al., 2014; Xu et al., 2014), maize (Tang et al., 2013), and cotton (Yu et al., 2011; Hulse-Kemp et al., 2015).'' Also more in that paper that might be useful. 
%Drosophila http://www.biorxiv.org/content/early/2017/02/06/106336
Segregation distortion provides putative evidence of haploid selection and can sometimes be attributed to meiotic drive and/or gametic competition \citep{Lalanne2004,Fishman2005,Leppala2008,Leppala2013,Didion2015,Didion2016}.
Where it has been characterized, meiotic drive generally occurs either during the production of male or female gametes only \citep{Ubeda:2005gw,Lindholm:2016cw}.
Gametic competition is also typically sex specific, occurring primarily among male gametes, because there are typically many more pollen/sperm than required for fertilization.
Gametic competition may be particularly common in plants, in which 60-70\% of all genes are expressed in the male gametophyte and these genes exhibit stronger signatures of selection than random genes \citep{Borg:2009jpa,Arunkumar:2013iq,Gossmann:2014dua}.
In addition, artificial selection pressures applied to male gametophytes are known to cause a response to selection \citep[e.g.,][]{Hormaza:1996cv,Ravikumar:2003uo,Hedhly:2004iv,Clarke:2004ir}. 
A smaller proportion of genes are thought to be expressed and selected during competition in animal sperm, although precise estimates are uncertain \citep{Zheng:2001fi,JOSEPH:2004haa,Vibranovski:2010et}. 
Recent studies have demonstrated that sperm competition in animals can alter haploid allele frequencies and increase offspring fitness \citep{Immler:2014im,Alavioon2017}.

There are various ways by which genes experiencing haploid selection could influence transitions between sex-determination systems. 
If we assume that haploid selection at any particular locus predominantly occurs in one sex (e.g., meiotic drive during spermatogenesis), then such loci experience a form of sex-specific selection. 
In this respect, we might expect that haploid selection would affect transitions between sex-determination systems in a similar manner to sex-specific diploid selection \citep[as explored by][]{vanDoorn:2007eu,vanDoorn:2010hu}. 
That is, new masculinizing mutations (neo-Y chromosomes) could be favoured via associations with alleles that are beneficial in the male haploid stage. 
On the other hand, sex ratios can also become biased by linkage between the sex-determining region and a locus that harbours genetic variation in haploid fitness. 
For example, there are several known cases of sex-ratio bias caused by sex-linked meiotic drive alleles \citep[][Chapter 3]{Burt:2006} or selection among X- and Y-bearing pollen \citep{Lloyd:1974tz,Conn:1981uw,Stehlik:2005ul,Stehlik:2006to,Field:2012fd,Field:2013cc}. 
It is not immediately clear how the spread of new sex-determination systems would be influenced by the combination of sex-ratio biases and associations between haploid selected loci and sex-determining regions.

@@

We find that sex-ratio biases caused by haploid selection can exert Fisherian sex-ratio selection upon novel sex-determiners but that their spread is also determined by selection on genetically-associated alleles. 
%Indeed, it is only when haploid-selected loci are tightly linked to the ancestral sex-determining region (and so sex-ratio biases are initially large) that we see an asymmetry between selection for XY to ZW transitions and ZW to XY transitions (e.g., because haploid selection in males only causes biased zygotic sex ratios in an ancestrally XY system).
Consequently, Fisherian sex ratio selection does not dominate and it is possible for selection on linked alleles to drive turnover between sex-determining systems despite causing increases in sex-ratio bias. 
In addition to considering haploid selection, another novel development in our model is that we consider loci that are in very tight linkage with the ancestral sex-determining region.
We show that transitions between male and female heterogamety can then evolve despite the fact that the neo-sex-determining locus is less closely linked to a locus under selection and therefore disrupts favourable ancestral associations between sex and the alleles favoured in that sex.
%Such transitions are not favoured in models lacking tight linkage and/or haploid selection. 
%Secondly, we allow sex-specific haploid selection to occur on a locus in tight or loose linkage with the ancestral sex-determining region. 
%In addition, when linkage between the ancestral sex-determining locus and a locus under haploid selection is tight, sex ratio biases can be strong such that there is an asymmetry between invasion of ancestrally XY and ancestrally ZW systems because, e.g., haploid selection in males only causes biased zygotic sex ratios in an ancestrally XY system.

@@

%\textcolor{blue}{
Neo-W alleles, on the other hand, can invade an ancestral XY system under some conditions (the full invasion conditions are given in the appendix; equations \ref{Binvasion} and \ref{Ainvasion}). 
That is, selection on loci within the non-recombining region of the SDR can favour the invasion of a less closely linked neo-W (Figure \ref{fig:SexAntagTighter}). 
%Although haploid selection can favour neo-W alleles because the ancestral sex ratio becomes male biased, this is not the only circumstance in which less tightly linked neo-W alleles invade. 
%\textcolor{red}{Sally: unpack, give one condition and relate to figures}
%For example, we note that it is possible for both neo-W haplotypes to spread ($\lambda_{mA}>1$ and $\lambda_{ma}>1$), in which case neo-W invasion can occur regardless of linkage to the selected locus. 
%For example, unlinked neo-W alleles can invade in the absence of any haploid selection. 
%This result is unexpected given the results of \cite{vanDoorn:2010hu}, who did not explicitly calculate equilibrium allele frequencies under tight linkage and generally concluded that heterogametic transitions can only occur when neo-sex-determining alleles are in tighter linkage with loci under sex-specific diploid selection. 
In fact, with tight linkage between the ancestral SDR and the selected locus, haploid selection and/or overdominance can favour completely unlinked neo-W alleles ($R=1/2$), allowing autosomes to become new sex chromosomes.
To develop an intuition for how less closely linked neo-W alleles invade, we first focus on cases where there is no haploid selection and discuss the additional effect of haploid selection in the appendix. 
%}
%Briefly, a neo-W can spread when the females it produces have higher fitness than the females produced under the ancestral SDR, which all have genotype XX. X chromosomes are sometimes found in males and therefore allele frequencies on the X can be suboptimal for females. In addition, the neo-W produces females from matings with Y-bearing male gametes. The allele associated with the Y can also be beneficial in females. 
%the neo-W invades under the ancestral XY sex-determining system, the X chromosome may not be able to as specialised on a female-beneficial allele

@@

%Haploid selection impacts the spread of neo-W haplotypes through its direct selective effect on females and their gametes, its indirect effect on the marginal fitness of alleles in females resulting from haploid selection in male gametes, and its effect on the sex ratio.
%These impacts expand the scenarios under which neo-W haplotypes can spread.
%For instance, overdominance is no longer required for neo-W-$a$ haplotypes to spread when there is haploid selection (e.g., when $a$ is female-beneficial, $A$ is male-beneficial, and meiotic drive in males favours $a$, Figure \ref{fig:regionMaleDrive}B).

@@

Significantly, equation \eqref{eq:lambda_ESD} is the same whether ESD is invading an ancestrally XY or ZW system (because $\lambda_{Y',XY} = \lambda_{W',ZW}$ and $\lambda_{W',XY} = \lambda_{Y',ZW}$).
Thus, Fisherian sex-ratio selection alone does not explain the invasion of ESD under weak selection because the sex ratio is only biased by male haploid selection when the ancestral sex-determination system is XY. 
Specifically, with male haploid selection, the neo-ESD is equally likely to invade when it equalizes the zygotic sex ratio (through $\lambda_{W',XY}$) and when it doesn't (through $\lambda_{Y',ZW}$). 
%This is again because haploid selection can facilitate the turnover of sex chromosomes either through sex ratio biases or through associations between the new sex determining region with selected alleles. 
%With weak selection, these two forces are nearly equal in strength. 
In addition, we note that ESD may not invade, even if the sex ratio is initially biased (e.g., with drive in males only, $r<1/2$, $h^\female=h^\male$, and $s^\female s^\male<0$, such that $\lambda_{W',XY}<1$, see Table \ref{tab:specialcases}).

@@

%\textcolor{red}{check last condition and the stability condition below are correct}
%\textcolor{blue}{The last condition looks good to me, although in your Turnover-norec-MFS.nb you look at YA fixed, so you have to flip everything (so I made Turnover-norec-MFS-MMO.nb to do this).
%The one issue I can find here is that you can also prevent $\lambda>1$ when the slope and intercept of the quadratic at $\lambda=1$ are negative (you only looked at both being positive). In this case we need $4w_{A}^{\female}w_{AA}^{\female}<(1-\alpha_\Delta^{\female})w_{a}^{\female}w_{Aa}^{\female}$, which also prevents $\psi>0$.
%I've added this in.
%It could also be the case that the slope and intercept are the same sign but the roots are imaginary - but this is never the case here.
%Stability condition below looks good to me (from matt version of turnoverSOM-MIKE.nb).}

@@

Still considering $w_{Aa}^\female>w_{AA}^\female$, the neo-W can also spread alongside the $a$ allele ($\lambda_{ma}>1$) if $w_{aa}^\female$ is large enough such that $(w_{Aa}^\female+w_{aa}^\female)/2>w_{AA}^\female$.
%there is sufficiently strong underdominance in females ($w_{aa}^\female>w_{Aa}^\female$) \textcolor{blue}{[this is describing directional selection, not overdominance - check the conditions for the case we want to talk about]}, such that $(w_{Aa}^\female+w_{aa}^\female)/2>w_{AA}^\female$.  
This can occur with overdominance or directional selection for $a$ in females (Figure \ref{fig:regionplots}B,C).
In this case, $a$ is favoured in females (comparing $Aa$ to $AA$ genotypes in females) but $A$ is fixed on the X due to selection in males. 
The neo-W-$a$ haplotype can spread because it produces females with higher fitness $Aa$ and $aa$ genotypes.

@@

\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=1.5\linewidth]{PositionPlot_SexAntagTighter_MaleDrive_Mike}}
\caption{
When there is sexually-antagonistic selection and haploid selection, a neo-W may invade for any $R$.
Panel A shows that the invasion fitness of a neo-W is positive where linkage is tight, even when $r<R$ (unshaded region). \textcolor{red}{(remove A?)}
In panel B, we vary the recombination rate between the neo-W and the selected locus ($R$) for a fixed recombination rate between the ancestral-SDR and the selected locus ($r=0.005$).
Coloured markers show recombination rates for which the temporal dynamics of neo-W invasion are plotted in panel C (black $R=0.001$, red $R=0.02$, blue $R=0.1$, green $R=0.5$). 
The diploid selection parameters used in this plot are the same as in Figure \ref{fig:SexAntagTighter}, marked by an asterisk in Figure \ref{fig:regionMaleDrive}A: $w_{AA}^\female=1.05$, $w_{Aa}^\Hermaphrodite=1$, $w_{aa}^\female=0.85$, $w_{AA}^\male=0.85$, $w_{aa}^\male=1.05$, $\alpha_{\Delta}^\male=-0.08$, except that there is also male meiotic drive in favour of the $a$ allele, $\alpha_{\Delta}^\male=-0.08$.
When $R=0.5$ (green curve), the neo-W does not reach fixation and X,Y,Z, and W alleles are all maintained in the population, see Figure \ref{fig:freqAll}C.
}
\label{fig:SexAntagTighterMaleDrive}
\end{figure}

\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=\linewidth]{Region_Plot_combined_MaleDrive_Mike}}
\caption{
Meiotic drive in males affects whether neo-W-$A$ and neo-W-$a$ haplotypes spread when the ancestral-XY locus is tightly linked to a locus under selection ($r=0$). 
We vary the fitness of male homozygotes relative to heterozygotes ($w_{Aa}^\Hermaphrodite=1$) and only consider stable equilibria at which both \textbf{A} locus allele are maintained and the $a$ allele is initially fixed on the Y, region outlined. 
In panels A-C, meiotic drive in males favours the $a$ allele ($\alpha_{\Delta}^\male=-0.16$), creating male-biased sex ratios and generally increasing $\lambda_{WA}$ and $\lambda_{Wa}$. 
By contrast, $\lambda_{WA}$ and $\lambda_{Wa}$ tend to be reduced when meiotic drive in males favours the $A$ allele ($\alpha_{\Delta}^\male=0.16$), panels D-F. 
We consider three forms of selection in females: directional selection in favour of the $A$ allele (panels A and D, $w_{aa}^\female=0.85$, $w_{AA}^\female=1.05$), direction selection in favour of the $a$ allele (panels B and E, $w_{aa}^\female=1.05$, $w_{AA}^\female=0.85$), and overdominance (panels C and F, $w_{aa}^\female=0.6$, $w_{AA}^\female=0.6$).
%\textcolor{red}{Panel F mislabelled, should have $\lambda_{Wa}>1$ \& $\lambda_{WA}<1$ as the upper label that has the line.}
}
\label{fig:regionMaleDrive}
\end{figure}

\newpage
\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=\linewidth]{Region_Plot_combined_MaleGS_Mike}}
\caption{
Parameters for which neo-W-$A$ and neo-W-$a$ haplotypes spread when there is male gametic competition at a locus that is tightly linked to the ancestral-XY locus.
Diploid selection parameters ($w_{ij}^\Hermaphrodite$) are the same as those in Figure \ref{fig:regionMaleDrive}. 
The $a$ allele is favoured during male gametic competition in Panels A-C ($w_{a}^\male=1.16$, $w_{A}^\male=1$), which creates male biased sex-ratios and increases $\lambda_{WA}$ and $\lambda_{Wa}$. 
On the other hand, the $A$ allele is favoured during male gametic competition in Panels D-F ($w_{a}^\male=1$, $w_{A}^\male=1.16$) and $\lambda_{WA}$ and $\lambda_{Wa}$ tend to be reduced. 
Compared to the meiotic drive parameters in Figure \ref{fig:regionMaleDrive}, the effect of these male gametic competition parameters on the sex ratio is smaller. 
For example, in Figure \ref{fig:regionMaleDrive}A-C, the ancestral sex ratio is $\alpha^\male=0.58$ at equilibrium (B) and in panels A-C of this plot, the ancestral sex ratio is $w_{a}^\male/(w_{A}^\male+w_{a}^\male)=0.537$ at equilibrium (B). 
}
\label{fig:regionMaleGS}
\end{figure}

\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=\linewidth]{Region_Plot_combined_FemaleDrive_Mike}}
\caption{
Parameters for which neo-W-$A$ and neo-W-$a$ haplotypes spread when there is female meiotic drive at a locus that is tightly linked to the ancestral-XY locus.
Diploid selection parameters ($w_{ij}^\Hermaphrodite$) are the same as those in Figure \ref{fig:regionMaleDrive} and \ref{fig:regionMaleGS}. 
The $a$ allele is favoured by meiotic drive in females in Panels A-C ($\alpha_{\Delta}^\female=-0.16$), which increases $\lambda_{Wa}$ and decreases $\lambda_{WA}$. 
Female meiotic drive in favour of the $A$ allele (panels D-F, $\alpha_{\Delta}^\female=-0.16$) has the opposite effect. 
}
\label{fig:regionFemaleDrive}
\end{figure}

\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=\linewidth]{Region_Plot_combined_FemaleGS_Mike}}
\caption{
Parameters for which neo-W-$A$ and neo-W-$a$ haplotypes spread when there is female meiotic drive at a locus that is tightly linked to the ancestral-XY locus.
Diploid selection parameters ($w_{ij}^\Hermaphrodite$) are the same as those in Figure \ref{fig:regionMaleDrive}, \ref{fig:regionMaleGS}, and \ref{fig:regionFemaleDrive}. 
The $a$ allele is favoured during female gametic competition in females in Panels A-C ($\alpha_{\Delta}^\female=-0.16$), which increases $\lambda_{Wa}$ and decreases $\lambda_{WA}$. 
The $A$ allele is favoured during gametic competition in panels D-F ($\alpha_{\Delta}^\female=-0.16$), giving the opposite effect on $\lambda_{Wa}$ and $\lambda_{WA}$. 
}
\label{fig:regionFemaleGS}
\end{figure}

@@

\begin{landscape}
\begin{figure}[!h]
\centering
\centerline{
\includegraphics[width=\linewidth]{Freq_plot_combined_PloidAntag_Matt}
}
\caption{
Dynamics of sex-determining alleles during invasion by a neo-W allele. 
The curves show the frequencies of the neo-W (red), ancestral-Y (Blue), and $A$ allele among female gametes (solid curves) and among male gametes (dashed curves). 
In panel A, there is a complete transition from XY sex determination (XX-ZZ females and XY-ZZ males) to ZW sex determination (YY-ZW females and YY-ZZ males).  
In panels B-D polymorphism is maintained at both the ancestral XY locus and the neo-ZW locus, such that there are males with genotypes XY-ZZ or YY-ZZ and females with genotypes XX-ZZ, XX-ZW, XY-ZW, or YY-ZW. 
In panel A, selection is ploidally antagonistic with drive in males (parameters as in the green curve in Figure \ref{fig:SexRatioBad}B).
In panel B, there is overdominance in both sexes (parameters as the green curve in Figure \ref{fig:temporalOverdominance}C).
In panel C, there is male meiotic drive and sexually-antagonistic selection in diploids (parameters as the green curve in Figure \ref{fig:regionMaleDrive}C).
\textcolor{red}{(remove D?)}
Panel D has the same parameters as the red curve in Figure \ref{fig:regionPloidAntag}F, except $r=0$ (ploidy-antagonism with pollen competition).
In all cases, the initial equilibrium frequency has $a$ near fixed on the Y.
}
\label{fig:freqAll}
\end{figure}
\end{landscape}

%\begin{figure}[!h]
%\centering
%\centerline{
%\includegraphics[width=\linewidth]{Sallys_R_plot.pdf}
%}
%\caption{
%\textcolor{blue}{is this the one?}
%}
%\label{fig:lambda_R}
%\end{figure}

%\newpage
%\textcolor{red}{
%Add Sally's figure showing lambda for small r near equil A versus near equil B. Add references to this figure to appendix where we discuss whether lambdas can be greater than 1 with sexually antagonistic selection.}
%\textcolor{blue}{not sure which one you are talking about, but see Figure \ref{fig:lambda_R} }
%\textcolor{red}{Let's remove this for now.}
